---
title: "160819 - Load Published NHP data for Blood Manuscript"
author: "Scott Furlan"
output: html_document
---


```{r, echo=FALSE, results='hide'}
remove(list=ls());
library(devtools)
library(dendextend)
library(Vennerable)
library(probedeeper)
library(XLConnect)
library(simpleaffy)
library(xtable)
library(sva)
library(made4)
library(ellipse)
library(ggplot2)
library(gplots)

###LOAD NHP DATA####Load Data
# -------------------------------------
# specify key
# -------------------------------------
date<-"20160819"
key<-""
# -------------------------------------

# -------------------------------------
# find relative directory on Server
# -------------------------------------
setwd("~")
ROOT_DIR<-"~/Dropbox (Kean Lab)/AWS/Scott"
# -------------------------------------
# specify paths and load functions
# -------------------------------------
DATA_DIR <- paste(ROOT_DIR, "/FinalAnalysis/Published/Scripts_BTGVHD/data", sep="") # SPECIFY HERE
DATA_DIR2 <- "~/Dropbox (Kean Lab)/LK NHP/Microarray/Data Files/CEL Files"
PROG_DIR <- paste(ROOT_DIR, "/FinalAnalysis/Published/Scripts_BTGVHD/prog", sep="")      # SPECIFY HERE
RES_DIR  <- paste(ROOT_DIR, "/FinalAnalysis/Published/Scripts_BTGVHD/res", sep="")      # SPECIFY HERE
source(file.path(PROG_DIR, "SFFunc.R"))
finaleset.nhp<-readRDS(file.path(DATA_DIR, "finaleset"))
unfilteredeset.nhp<-readRDS(file.path(DATA_DIR, "unfilteredeset"))

# -------------------------------------
# load color NHP
# -------------------------------------
color.wb<-loadWorkbook(file.path(DATA_DIR, "BreakthroughColor.xlsx"))
color.ws<-readWorksheet(color.wb, "BreakthroughColor", header=TRUE)
ColObj.nhp<-new("ColObj", assign=color.ws, classvec=as.factor(finaleset.nhp$NetworkGroup))
ColObj.nhp<-ColObjInit(ColObj.nhp, LD=70)
```

##Check Assignments
```{r, echo=F}
animal.table<-data.frame(ID=paste(finaleset.nhp@phenoData@data$ID, " Day-", finaleset.nhp@phenoData@data$Day, sep=""), Group=ColObj.nhp@classvec, ArrayTimePoint=finaleset.nhp@phenoData@data$Group, stringsAsFactors=F)
animal.table.split<-split(animal.table$ID, as.factor(animal.table$Group))
animal.table.split.df<-as.data.frame(t(do.call(rbind, lapply(animal.table.split, '[', 1:max(sapply(animal.table.split, length))))))
tli.table <- xtable(animal.table.split.df)
```

```{r, echo=FALSE, results='asis'}
print(tli.table, type="html", include.rownames=FALSE, booktabs=TRUE, floating=FALSE, scalebox=0.6)
```


###LIMMA
```{r, echo=FALSE}
PD<-new("PDObj", eset=unfilteredeset, ColObj=ColObj.nhp, LimmaObj=LimmaObjCreate(unfilteredeset, ColObj.nhp))
```

##Figure 2C Colors
```{r}
step1<-40
step2<-40
line<-vector()
fill<-vector()
line[1]<-"#931a1d"
fill[1]<-LightenDarkenColor(line[1], step1)
line[2]<-LightenDarkenColor(line[1], step2)
fill[2]<-LightenDarkenColor(line[2], step1)
line[3]<-LightenDarkenColor(line[2], step2)
fill[3]<-LightenDarkenColor(line[3], step1)
line[4]<-LightenDarkenColor(line[3], step2)
fill[4]<-LightenDarkenColor(line[4], step1)
pie(rep(1,4), col=fill, labels=fill)
pie(rep(1,4), col=line, labels=line)

step1<-40
step2<-30
lineb<-vector()
fillb<-vector()
start<-"#F7931D"
lineb[1]<-LightenDarkenColor(start, -50)
fillb[1]<-LightenDarkenColor(lineb[1], step1)
lineb[2]<-LightenDarkenColor(lineb[1], step2)
fillb[2]<-LightenDarkenColor(lineb[2], step1)
lineb[3]<-LightenDarkenColor(lineb[2], step2)
fillb[3]<-LightenDarkenColor(lineb[3], step1)
lineb[4]<-LightenDarkenColor(lineb[3], step2)
fillb[4]<-LightenDarkenColor(lineb[4], step1)
pie(rep(1,4), col=fillb, , labels=fillb)
pie(rep(1,4), col=lineb, , labels=lineb)
```


#Figure 3A
``` {r Survival, echo=FALSE}
sel<-ColObj.nhp@classvec=="Breakthrough" | ColObj.nhp@classvec=="Primary"
pcadata<-t(exprs(finaleset.nhp)[,sel])

rownames(pcadata)<-make.unique(rownames(pcadata))
pca1 <- dudi.pca(pcadata, scann = FALSE, nf=10)
s.class(dfxy = pca1$li, fac = ColObj.nhp@classvec[sel], xax = 1, yax = 3)
```

#Figure 3B
```{r}
pcadata<-t(exprs(finaleset.nhp))
#pcadata<-t(data.sel.nhp)
rownames(pcadata)<-make.unique(rownames(pcadata))
pca1 <- dudi.pca(pcadata, scann = FALSE, nf=10)

###PCA SURV###
selectionindex<-vector()
selectionindex[!is.na(suppressWarnings(as.numeric(finaleset.nhp$Survival)))]<-"TRUE"
selectionindex[ColObj.nhp@classvec=="Controlled"]<-"FALSE"
selectionindex[is.na(selectionindex)]<-"FALSE"
selectionindex<-as.logical(selectionindex)
Survival.sel<-as.numeric(finaleset.nhp$Survival[selectionindex])
lm.df<-data.frame(PC1=as.numeric(pca1$li$Axis1[selectionindex]), Survival=as.numeric(log(Survival.sel)), Group=as.factor(as.character(ColObj.nhp@classvec[selectionindex])))
reg1<-(lm(PC1~Survival, lm.df))
summary(reg1)
anova(reg1)
lm.df.melt<-melt(lm.df)
g<-ggplot()+
  geom_point(data=lm.df, aes(x=Survival, y=PC1), col=ColObj.nhp@full$line[selectionindex], size=3)+
  geom_smooth(data=lm.df, aes(x=Survival, y=PC1), method=lm)+
  theme_bw()+
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())+
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank())+
  labs(x="log Survival", y="PC1")
print(g)
```




#Figure 1C
```{r}
library(survival)
survobj<-Surv(Survival.sel)
survfit<-survfit(survobj~lm.df$Group)
plot(survfit, col=ColObj.nhp@match$line[c(2,5)])
print(survfit)
```

#Make GSEA Files
```{r}
###GSEA Data###
data.sel.nhp<-exprs(finaleset)
classvec.sel.nhp<-ColObj.nhp@classvec
data.gct.nhp<-data.sel.nhp[,order(classvec.sel.nhp)]
colnames(data.gct.nhp)<-make.unique(as.character(classvec.sel.nhp)[order(classvec.sel.nhp)])
gsea.write.gct(data.gct.nhp, file.path(DATA_DIR, "NHP_Allnetworkmodules.gct"))
classvec.sel.cls.nhp<-classvec.sel.nhp[order(classvec.sel.nhp)]
gsea.write.cls(as.factor(classvec.sel.cls.nhp), file.path(DATA_DIR, "NHP_Allnetworkmodules.cls"))
GSEA.ReadClsFile(file.path(DATA_DIR, "NHP_Allnetworkmodules.cls"))
```

#Figure 3C + 4A
####ssGSEA data were generated using the GenepPattern Server using the following input files
1. NHP_Allnetworkmodules.gct
2. NHP_Allnetworkmodules.cls

####Output file is ssGSEA_masterDB_nodupes.gct

####The ConstellationMapping code was run on GenePattern Server and had output details summarized in the following textfiles
1. PrimaryUpConstellationMap.txt
2. BreakthroughUpConstellationMap.txt
####Input files included
1. ssGSEA_masterDB_nodupes.gct
2. NHP_Allnetworkmodules.cls


#Figure 3D
```{r}
Comparison1<-"Primary"
Comparison2<-c("Healthy.Control", "Breakthrough")
masterDB<-readGMT.SF(file.path(DATA_DIR, "MSigDBALL.gmt"))
toGMT<-masterDB[grep("GSE14308",names(masterDB))]
MCGO3<-MultipleClassGSEA(data.gct.nhp,
                         Comparison1, Comparison2, toGMT, 
                         classvec=as.factor(classvec.sel.cls.nhp),
                         runGSEAcode=TRUE, reshuffling.type="gene.labels",
                         uniquelabel="GSE14308-rearranged-PvOthers",
                         directory=DATA_DIR)
```

#Figure4C
```{r}
MAboxplot("MKI67", PD, reorder=c(4,1,5,3,2))
MAboxplot("GZMB", PD, reorder=c(4,1,5,3,2))
```


##Figure 4C Pivot Transcripts
```{r}
limmaObj<- LimmaObjCreate(finaleset, ColObj.nhp, lfc.thresh=0.5)
slotNames(limmaObj)
str(limmaObj@Contrasts)
limmaObj@Contrasts$meaning
limmaObj2<- LimmaObjCreate(unfilteredeset, ColObj.nhp, lfc.thresh=0.5)

Symbol<-rownames(limmaObj@DEGenes[[14]])[rownames(limmaObj@DEGenes[[14]]) %in% rownames(limmaObj@DEGenes[[16]])]
colnames(limmaObj@DEGenes[[14]])
X<-limmaObj@DEGenes[[14]][Symbol,]$logFC
Y<-limmaObj@DEGenes[[16]][Symbol,]$logFC
t<--log(limmaObj@DEGenes[[14]][Symbol,]$adj.P.Val)
df<-data.frame(Symbol=Symbol, X=X, Y=Y, t=t)
ggplot(df)+
  geom_point(aes(x=X, y=Y, size=t))

Symbol<-rownames(limmaObj@AllGenes[[14]])[rownames(limmaObj@AllGenes[[14]]) %in% rownames(limmaObj@AllGenes[[16]])]
colnames(limmaObj@AllGenes[[14]])
X<-limmaObj@AllGenes[[14]][Symbol,]$logFC
Y<-limmaObj@AllGenes[[16]][Symbol,]$logFC
t<--log(limmaObj@AllGenes[[14]][Symbol,]$adj.P.Val)
df2<-data.frame(Symbol=Symbol, X=X, Y=Y, t=t)
Color<-vector()
Color[which(df2$Symbol %in% df$Symbol)]<-"red"
Color[which(!df2$Symbol %in% df$Symbol)]<-"black"
df2$Color<-Color

BTUP<-as.character(df$Symbol[intersect(which(df$X>0), which(df$Y<0))])
PUP<-as.character(df$Symbol[intersect(which(df$X<0), which(df$Y>0))])
df3<-df[df$Symbol %in% c(BTUP, PUP),]


ggplot(df2)+
  geom_point(aes(x=X, y=Y, size=t), colour="black", alpha=0.1)+
  geom_point(data=df, aes(x=X, y=Y, size=t), colour="blue", alpha=0.2)+
  geom_point(data=df3, aes(x=X, y=Y, size=t), colour="red")+
  geom_text(data=df3, aes(x=X, y=Y, label=Symbol), hjust=0, vjust=0, colour="red")+
  theme_bw()
```


##Figure 4D Hill Heatmap
###1. DE list of Breakthrough vs Auto vs transcripts listed in the Hill paper.
``` {r, echo=FALSE}
#Hill
con<-file(file.path(DATA_DIR, "Hill.txt"))
HILL_UP<-toupper(unlist(strsplit(readLines(con, warn="FALSE")[1], " ")))
HILL_DOWN<-toupper(unlist(strsplit(readLines(con, warn="FALSE")[2], " ")))
close(con)
Hill<-c(HILL_UP, HILL_DOWN)

df<-data.frame(Gene=rownames(PD@LimmaObj@AllGenes[[1]][Hill,]), BvA=PD@LimmaObj@AllGenes[[1]][Hill,1], BvHC=PD@LimmaObj@AllGenes[[14]][Hill,1])
mat<-as.matrix(df[,2:3])
rownames(mat)<-df$Gene
dir<-vector()
dir[1:12]<-df$BvHC[1:12]>0
dir[13:24]<-df$BvHC[13:24]<0
names(dir)<-as.character(df$Gene)
dir2<-vector()
dir2[1:12]<-df$BvA[1:12]>0
dir2[13:24]<-df$BvA[13:24]<0
names(dir2)<-as.character(df$Gene)
dir2

dir3<-dir2 | dir
rmat<-t(as.matrix(cbind(c("Black", "Grey")[as.factor(dir3)], RowSide)))
heatmap.3(mat, dendrogram="none", Rowv=FALSE, scale="none", col=gsea.col, RowSideColors=rmat, trace="none", density.info="none", cexCol = 2, main="RowSide by Auto OR HC")
```



#Appendix
```{r Appendix,echo=FALSE}
sessionInfo()
getwd()
```


